---
title: "Homework 3"
author: "Gauri Bhatkhande"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "botttom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")
```
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user/order variables -- and item variables. `r variable.names(instacart)`

How many aisles and which are most items from? 
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
Lets make a plot 

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
  )%>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Lets make a table 

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
    knitr::kable()
```
Apples vs ice cream 

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))%>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
pivot_wider(
  names_from = order_dow, 
  values_from = mean_hour
)
```

## Problem 2

Loading the dataset
```{r}
accelero_data = read_csv(file = "./Data/accel_data.csv")
```

Tidying and wrangling the dataset

```{r}
accelero_data = janitor::clean_names(accelero_data) %>%
                pivot_longer(
                   activity_1:activity_1440,
                   names_to = "activity_for_minute",
                   values_to = "activity_count") %>%
  separate(activity_for_minute, c("word", "minute_of_the_day"), "_") %>%
  select(-word) %>%
  mutate(
    minute_of_the_day = as.numeric(minute_of_the_day)
  ) %>%
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
        )%>%
  mutate(day_type = case_when(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "Weekday",
                              day %in% c("Saturday", "Sunday") ~ "Weekend")
        )
accelero_data
```
The variables in the data are `r variable.names(accelero_data)`. There are `r nrow(accelero_data)` rows (observations) and `r ncol(accelero_data)` columns in the dataset. The number of rows or observations signifies activity count for each minute of the day for each day of the week for 5 weeks (35 days). 

Making a table of total activity for each day of the week

```{r}
accelero_data %>%
  group_by(week, day) %>%
  summarize(total_activity = sum(activity_count)) %>%
 knitr::kable()
```
He tends to have more physical activity during the weekdays than the weekends for the first couple of weeks. Later his activity on weekends increases. However, during the Saturdays of the 4th and 5th weeks, he had barely any activity. They could have been rest days. 


```{r}
accelero_data %>%
ggplot(aes(x = minute_of_the_day, y = activity_count, color = day, )) + stat_smooth(se=FALSE) +
labs(
    title = "Activity time courses per day",
    x = "Minute of the day",
    y = "Activity count",
    caption = "Data from Advanced Cardiac Care Center of Columbia University Medical Center"
  ) +
  theme(legend.position = "bottom")
  
```

Except for Sunday, Most of the activity is higher during the latter half of the day. It appears that on Fridays the activity counts were the highest among all days for the later half of the day. On Sunday the activity is high towards the middle of the day. 


